\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{geometry}
\usepackage{titlesec}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{hyperref}

% Operational formatting
\geometry{a4paper, margin=1in}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.8em}
\linespread{1.1}

% Section styling
\titleformat{\section}{\Large\bfseries\sffamily}{\thesection}{1em}{}
\titleformat{\subsection}{\large\bfseries\sffamily}{\thesubsection}{1em}{}

\title{\textbf{\sffamily Social Capacity Praxis Manual}}
\author{Operational Framework}
\date{}

\begin{document}

\maketitle

\section*{Purpose}
This manual translates the Social Capacity theoretical framework into operational metrics, diagnostic tools, and intervention strategies. It is designed for researchers, policymakers, and organizers.

\section*{1. Operational Equations}

\subsection*{1.1 Outcome Realization}
Outcomes emerge from objective amplification before subjective valuation.
\begin{equation}
O_{im} = \sum_j X_{ij} A_{jm}
\end{equation}
\textbf{Metric:} $O_{im}$ is measurable (e.g., liters of water, hours of care, bits of bandwidth).

\subsection*{1.2 Sustainability Thresholds}
Systems exhibit bifurcation dynamics based on participation levels.
\begin{equation}
\frac{dA_j}{dt} = \alpha_j \left( \sum_i X_{ij} - \kappa_j \right)
\end{equation}
\begin{itemize}
    \item If $\sum X < \kappa_j$: System decays (commons tragedy).
    \item If $\sum X \ge \kappa_j$: System sustains or grows.
    \item $\kappa_j$: The maintenance cost (e.g., minimum maintainers for OSS, minimum tax base).
\end{itemize}

\subsection*{1.3 Inequality Structure}
Inequality is defined as the ratio of maximum accessible amplification.
\begin{equation}
I_{ik} = \frac{\max_j A_j^{(i)}}{\max_j A_j^{(k)}} \quad \text{assuming } X_{ij} = X_{kj}
\end{equation}
If $I_{ik} \gg 1$, agent $i$ has structural privilege. Inequality is not difference in effort ($X$), but difference in available multiplication ($A$).

\subsection*{1.4 Power Analysis}
Operational power is the capacity to shift others' allocations or system properties.
\begin{align}
P_{\text{alloc}} (i \to k) &= \sum_j \left| \frac{\partial X_{kj}}{\partial X_{ij}} \right| \quad \text{(Influence)} \\
P_{\text{amp}} (i \to j) &= \frac{\partial A_j}{\partial X_{ij}} \quad \text{(Control)}
\end{align}
\textit{Diagnosis:} A "boss" is an agent where $\frac{\partial X_{kj}}{\partial X_{ij}} \approx 1$ (k moves where i directs).

\section*{2. Measurement Protocol}

To measure $A_{jm}$ (Amplification) empirically:

\subsection*{2.1 Difference-in-Differences}
\begin{equation}
\hat{A}_{jm} = \frac{\mathbb{E}[O_{im} \mid X_{ij}=1, Z] - \mathbb{E}[O_{im} \mid X_{ij}=0, Z]}{\mathbb{E}[O_{im} \mid \text{baseline}]}
\end{equation}
Where $Z$ is the vector of confounders (e.g., prior wealth, education).

\section*{3. Diagnostic Checklist}

Use this checklist to analyze any social system (e.g., a coop, a corporation, a city).

\begin{enumerate}
    \item \textbf{Identify Goals ($W$)}: What does the community actually value? (Health, autonomy, profit?)
    \item \textbf{Map Systems ($j$)}: What distinct relations/institutions exist?
    \item \textbf{Estimate Amplification ($A_j$)}:
        \begin{itemize}
            \item Does System J multiply effort or just extract it?
            \item Is $A_j$ visible to participants? (Transparency $\eta$)
        \end{itemize}
    \item \textbf{Check Sustainability ($\kappa$)}:
        \begin{itemize}
            \item Is $\sum X_{ij} \ge \kappa_j$?
            \item Are there free riders ($A > 0, X=0$)?
        \end{itemize}
    \item \textbf{Analyze Barriers ($\Delta X$)}:
        \begin{itemize}
            \item Can members leave? (Switching costs)
            \item Do they know about better alternatives? (Information costs)
        \end{itemize}
\end{enumerate}

\section*{4. Intervention Strategy}

To increase freedom (Correction Velocity $V_i$):

1.  **Increase Transparency ($\eta$)**: Make $A_j$ distinct and measurable.
2.  **Reduce Switching Costs**: Guarantee basic needs so $X_{ij}$ isn't forced by survival.
3.  **Lower Sustainability Thresholds ($\kappa$)**: Automate maintenance to make commons easier to sustain.

\end{document}
